mouse_microbiome

### 2018-07-06 Manuscript_Mouse_Microbiome.Rmd
more figs

```{r setup, include = FALSE}

source('C://Users//John//Documents//initialize_manuscript.R')

@@

@@

#fix data entry problem1
filt <- x$Mouse == 9 & x$Hour == 168
x$Cohoused[filt] <- 'Y'
x$Pair[filt] <- 2
x$Abx[filt] <- 'N'
x$Dose[filt] <- 'No'

#fix data entry problem2
filt <- x$Mouse == 7 & x$Hour == 168
x$Cohoused[filt] <- 'Y'
x$Pair[filt] <- 4
x$Abx[filt] <- 'Y'
x$Dose[filt] <- 'Recovery'

@@

p1 <- j %>% 
  filter(!is.na(Pair)) %>%
  mutate(Pair = paste('Cage', Pair)) %>%
  ggplot(aes(x = PCoA1, y = PCoA2, color = Abx, fill = Abx)) + 
  geom_point(aes(alpha = Hour), shape = 21, size = 4) + 
  geom_point(shape = 21, size = 4, fill = NA) + 
  geom_path() + 
  geom_text(aes(label = label), color = 'black', na.rm=TRUE) +
  facet_grid(Pair~Abx) + th

@@

@@

**(ref:figPCoAPaired):** Figure legend.

@@

p2 <- j %>% 
  filter(is.na(Pair)) %>%
  group_by(Abx) %>%
  mutate(Pair2 = as.numeric(as.factor(Mouse))) %>%
  ggplot(aes(x = PCoA1, y = PCoA2, color = Abx, fill = Abx)) + 
  geom_point(aes(alpha = Hour), shape = 21, size = 4) + 
  geom_point(shape = 21, size = 4, fill = NA) + 
  geom_path() + 
  geom_text(aes(label = label), color = 'black', na.rm=TRUE) +
  facet_grid(Pair2~Abx) + th

@@

@@

**(ref:figPCoASolitary):** Figure legend.

@@

j <- tax %>% 
  gather(tax_level, name, -otu) %>% 
  full_join(x, by = 'otu') %>% 
  filter(tax_level != 'Domain') %>% 
  filter(abun > 0)

j <- j %>% 
  mutate(tax_level = factor(tax_level, levels = c('Phylum','Class','Order','Family','Genus','Species'))) %>%
  group_by(tax_level) %>% 
  mutate(total_rich = n_distinct(name)) %>% 
  group_by(Mouse, tax_level) %>% 
  mutate(rich = n_distinct(name), val = rich / total_rich)

p1 <- j %>% 
  ggplot(aes(x = tax_level, y = val, group = Mouse)) + 
  geom_line() + 
  expand_limits(y = c(0,1)) +
  th

@@

**(ref:figTaxaSharedAcrossMice):** Percent of total phyla present within single mice, grouped by taxonomic level. Need to add OTU as a final category. First, I categorize each OTU in each antibiotic treated mouse into *Absent* (never was there), *Extinct* (was there at the end of antibiotic treatment but then disappeared), *Persistent* (was present throughout the experiment), and *Immigrant* (Was absent at the end of AB treatment, but colonized during recover).

The first plot shows the Hour at which immigrants arrive, plotted by the abundance of that immigrant in the control mouse.

@@

#I am assuming that OTUs that are absent from a given mouse in the last two samples collected during antibiotic dosing (i.e., at hours 73 and 97) are not in the mouse. This window can be tweaked.

j <- x %>%
  group_by(Mouse) %>%
  filter(Hour >= 63 | Abx == 'N') %>%
  group_by(Mouse, otu) %>%
  mutate(
    abun0 = sum(abun[Hour <= 97]),
    status0 = if (sum(abun[Hour <= 97]) > 0) 'present' else 'absent',
    status1 = if (sum(abun[Hour > 97]) > 0) 'present' else 'absent',
    status = ifelse(status0 == 'absent' & status1 == 'absent', 'absent',
                    ifelse(status0 == 'absent' & status1 == 'present', 'immigrant',
                           ifelse(status0 == 'present' & status1 == 'absent', 'extinct',
                                  'persistent'))),
    arrival_time = if (status[1] == 'immigrant') min(Hour[Hour > 97 & abun > 0]) else NA) %>%
  ungroup()

tmp <- j %>%
  filter(!is.na(Pair)) %>%
  group_by(Pair, otu) %>%
  mutate(control_abun = ifelse(Abx == 'Y', unique(abun0[Abx == 'N']), 0))

p <- tmp %>% 
  filter(status == 'immigrant') %>%
  #filter(control_abun > 0) %>%
  mutate(
    time_to_arrival = arrival_time - 120,
    control_abun = control_abun * 100 + 0.01) %>%
  ggplot(aes(x = control_abun, y = time_to_arrival)) + 
    geom_jitter(alpha = 0.05, color = 'red') + 
    scale_x_log10() + 
    stat_smooth(color = 'black', method = 'lm') +
    stat_smooth(color = 'black', lty = 3) +
    labs(x = "OTU abundance in cohoused control mouse + 0.1 %", y = 'Hours before OTU arrival in Abx treated mouse') +
    th

@@

@@

**(ref:figPropPressure):** Figure legend.

@@

#numbers of OTUs, colored by local status
tmp %>% 
  filter(Abx == 'Y' & status %in% c('immigrant','persistent') & abun > 0) %>% 
  mutate(status = as.character(status)) %>%
  ungroup() %>% 
  arrange(status) %>% 
  ggplot(aes(x = factor(Hour), fill = status)) + 
    geom_bar() + 
    facet_wrap(~Pair, ncol = 1) +
    labs(x = "Hour", y = "Number of OTUs") +
    th

#Relative contribution of OTUs, grouped by local status
p <- tmp %>%
  filter(Abx == 'Y' & status %in% c('immigrant','persistent')) %>% 
  mutate(status = as.character(status)) %>%
  ungroup() %>% 
  arrange(status) %>% 
  ggplot(aes(x = factor(Hour), y = abun, fill = status)) + 
    geom_bar(stat = 'identity') + 
    facet_wrap(~Pair, ncol = 1) + 
    labs(x = "Hour", y = "Relative abundances") +
    th

plot_it(p, name = 'figImmAbunTime', dir = image_dir, height = 7, width = 5)

@@

(ref:figImmAbunTime) Figure `r refs$figImmAbunTime`

`r pagebreak` ![](images\\figImmAbunTime.png)

**(ref:figImmAbunTime):** Figure legend.

@@

p <- x %>%
  left_join(tax) %>%
  filter(otu %in% j$otu) %>%
  mutate(label = paste(Family, 'OTU', as.numeric(as.factor(otu)))) %>%
  left_join(j) %>%
  ggplot(aes(x = Hour, y = abun + 1, color = Abx)) + 
  annotate(geom = 'rect', xmin = 1, xmax = 120, ymin = 0, ymax = Inf, alpha = 0.3, fill = 'red') + 
  geom_point() + 
  stat_smooth() +
  #geom_text(label = "Antibiotics", x = 60, aes(y = max(abun))) +
  #geom_text(label = "Recovery", x = 216, aes(y = max(abun))) +
  scale_y_log10() +
  scale_color_manual(values = c('black','red')) +
  facet_wrap(~label) + 
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.background = element_blank(),
    legend.position = 'bottom')

@@

@@

`r pagebreak` ![](images\\figImmAbunTime.png)

**(ref:figMostAffectedOTUs):** Figure legend.

>>>>

#initialize manuscript formatting script
source('C://Users//John//Documents//initialize_manuscript.R')

@@

`r pagebreak`

![](images/setup_Reese2018.png)

`r ref('figure', 'figReese2018Setup')`

(ref:figReese2018Setup) Figure `r refs$figReese2018Setup`

**(ref:figReese2018Setup):** A figure drawn from @Reese2018 showing the experimental setup and data collection schedule.

`r pagebreak`

@@

#fix data entry problems
x$Cohoused[x$Mouse == 3 & x$Hour == 168] <- 'N'
x$Cohoused[x$Mouse == 6 & x$Hour == 168] <- 'N'
x$Cohoused[x$Mouse == 7 & x$Hour == 168] <- 'Y'
x$Pair[x$Mouse == 7 & x$Hour == 168] <- 4
x$Abx[x$Mouse == 7 & x$Hour == 168] <- 'Y'
x$Dose[x$Mouse == 7 & x$Hour == 168] <- 'Recovery'
x$Cohoused[x$Mouse == 9 & x$Hour == 168] <- 'Y'
x$Pair[x$Mouse == 9 & x$Hour == 168] <- 2
x$Abx[x$Mouse == 9 & x$Hour == 168] <- 'N'
x$Dose[x$Mouse == 9 & x$Hour == 168] <- 'No'
x$Cohoused[x$Mouse == 12 & x$Hour == 168] <- 'N'
x$Cohoused[x$Mouse == 16 & x$Hour == 168] <- 'N'

@@

p1 <- j %>% 
  filter(!is.na(Pair)) %>%
  mutate(Pair = paste('Cage', Pair)) %>%
  ggplot(aes(x = PCoA1, y = PCoA2, color = Abx, fill = Abx)) + 
  geom_point(aes(alpha = Hour), shape = 21, size = 4) + 
  geom_point(shape = 21, size = 4, fill = NA) + 
  geom_path() + 
  geom_text(aes(label = label), color = 'black', na.rm=TRUE) +
  scale_color_manual(values = c('black','red')) +
  scale_fill_manual(values = c('black','red')) +
  facet_grid(Pair~Abx) + 
  th

@@

p1

@@

**(ref:figPCoAPaired):** PCoA of mouse microbiomes over time, during and after 120 hours of antibiotic treatment. Control mice (left) and treated mice (right) were cohoused after treatment (hour 120) for the duration of the experiment.

@@

p2 <- j %>% 
  filter(is.na(Pair)) %>%
  group_by(Abx) %>%
  mutate(Pair2 = as.numeric(as.factor(Mouse))) %>%
  ggplot(aes(x = PCoA1, y = PCoA2, color = Abx, fill = Abx)) + 
  geom_point(aes(alpha = Hour), shape = 21, size = 4) + 
  geom_point(shape = 21, size = 4, fill = NA) + 
  geom_path() + 
  geom_text(aes(label = label), color = 'black', na.rm=TRUE) +
  scale_color_manual(values = c('black','red')) +
  scale_fill_manual(values = c('black','red')) +
  facet_grid(Pair2~Abx) + th

@@

p2

@@

**(ref:figPCoASolitary):** PCoA of mouse microbiomes over time, during and after 120 hours of antibiotic treatment. Control mice (left) and treated mice (right) were not cohoused at any point during the experiment.

@@

j <- tax %>% 
  gather(tax_level, name, -otu) %>% 
  full_join(x, by = 'otu') %>% 
  filter(Hour <= 120) %>%
  filter(tax_level != 'Domain') %>% 
  filter(abun > 0)

j <- j %>% 
  mutate(tax_level = factor(tax_level, levels = c('Phylum','Class','Order','Family','Genus','Species'))) %>%
  group_by(tax_level) %>% 
  mutate(total_rich = n_distinct(name)) %>% 
  group_by(Mouse, tax_level) %>% 
  mutate(rich = n_distinct(name), val = rich / total_rich,
         Abx = ifelse(Abx == 'Y', 'Abx', 'No abx'))

p1 <- j %>% 
  ggplot(aes(x = tax_level, y = val, group = Mouse, color = Abx)) + 
  #geom_line() + 
  stat_summary(fun.data = "mean_sdl", fun.args = list(mult=1), aes(group = Abx), position = position_dodge(width = 0.2)) +
  scale_color_manual(values = c('black','red')) +
  expand_limits(y = c(0,1)) +
  labs(x = 'Taxonomic level', y = 'Percent of all Phyla present across all mice')
  th

@@

**(ref:figTaxaSharedAcrossMice):** Percent of phyla observed across all mice present within single mice, assessed at increasingly specific taxonomic levels.

@@

j <- x %>%
  group_by(Mouse, Hour) %>%
  mutate(rich = length(unique(otu[abun > 0]))) %>%
  group_by(Mouse, Hour, Abx, Cohoused, Dose, Pair, rich) %>%
  do(shannon = diversity(spread(select(., otu, abun), otu, abun))) %>%
  mutate(shannon_val = shannon[[1]])

#must fix screwy data!!! (when I paste mouse, abx, cohoused, there are some unexpected combinations)
j <- j %>%
  group_by(Mouse, Abx, Cohoused) %>%
  filter(length(shannon_val) > 1) %>%
  ungroup() %>%
  mutate(
    evenness = shannon_val / log(rich),
    Abx = ifelse(Abx == 'Y', 'Abx', 'No abx'),
    Mouse = ifelse(Cohoused == 'Y', paste(Abx, "Cohoused", Mouse), paste(Abx, 'Solitary', Mouse)))

p <- j %>%
  ggplot(aes(x = Hour, y = shannon_val, color = Abx)) +
    geom_point() +
    stat_smooth() +
    scale_color_manual(values = c('red','black')) +
    geom_rect(aes(xmin = 0, xmax = 120, ymin = -Inf, ymax = Inf), alpha = 0.3, fill = 'red', color = NA, data = j %>% group_by(Mouse) %>% filter(row_number() == 1 & Abx == 'Abx')) + 
    facet_wrap(~Mouse, ncol = 3) +
    labs(x = 'Hour', y = 'Shannon diversity') +
    th

plot_it(p, name = 'figShannonOverTime', dir = image_dir, width = 6, height = 8)

p

(ref:figShannonOverTime) Figure `r refs$figShannonOverTime`

`r pagebreak` ![](images\\figShannonOverTime.png)

**(ref:figShannonOverTime):** Shannon diversity over time within mice.

p <- j %>%
  ggplot(aes(x = Hour, y = rich, color = Abx)) +
    geom_point() +
    stat_smooth() +
    scale_color_manual(values = c('red','black')) +
    geom_rect(aes(xmin = 0, xmax = 120, ymin = -Inf, ymax = Inf), alpha = 0.3, fill = 'red', color = NA, data = j %>% group_by(Mouse) %>% filter(row_number() == 1 & Abx == 'Abx')) + 
    facet_wrap(~Mouse, ncol = 3) +
    labs(x = 'Hour', y = 'OTU richness') +
    th

plot_it(p, name = 'figRichOverTime', dir = image_dir, width = 6, height = 8)

p

(ref:figRichOverTime) Figure `r refs$figRichOverTime`

`r pagebreak` ![](images\\figRichOverTime.png)

**(ref:figRichOverTime):** OTU richness over time within mice.

p <- j %>%
  ggplot(aes(x = Hour, y = evenness, color = Abx)) +
    geom_point() +
    stat_smooth() +
    scale_color_manual(values = c('red','black')) +
    geom_rect(aes(xmin = 0, xmax = 120, ymin = -Inf, ymax = Inf), alpha = 0.3, fill = 'red', color = NA, data = j %>% group_by(Mouse) %>% filter(row_number() == 1 & Abx == 'Abx')) + 
    facet_wrap(~Mouse, ncol = 3) +
    labs(x = 'Hour', y = 'Community evenness') +
    th

plot_it(p, name = 'figEvennessOverTime', dir = image_dir, width = 6, height = 8)

p

(ref:figEvennessOverTime) Figure `r refs$figEvennessOverTime`

`r pagebreak` ![](images\\figEvennessOverTime.png)

**(ref:figEvennessOverTime):** OTU community evenness (i.e., equitability, i.e., Shannon diversity H / log(OTU richness))

#I am assuming that OTUs that are absent from a given mouse in the last three samples collected during antibiotic dosing (i.e., at hours 63, 73 and 97) are not in the mouse. This window can be tweaked.

#set the window. The edges are included.
window0 <- 63 
window1 <- 120

j <- x %>%
  group_by(Mouse, Abx, Pair, otu) %>%
  summarise(
    abunPre = if (unique(Abx) == 'Y') sum(abun[Hour <= window0]) else NA,
    abun0 = if (unique(Abx) == 'Y') sum(abun[Hour >= window0 & Hour <= window1]) else sum(abun[Hour < window1]),
    abun1 = sum(abun[Hour > window1]),
    statusAB = if (!is.na(abunPre) & abunPre == 0 & abun0 == 0) 'absent' else 'persistent',
    statusAB = if (!is.na(abunPre) & abunPre > 0 & abun0 == 0) 'extinct' else statusAB,
    ETD = if (statusAB == 'extinct') max(Hour[Hour <= window0 & abun > 0]) else NA,
    ETD = if (statusAB == 'persistent') Inf else ETD,
    status0 = if (abun0 > 0) 'present' else 'absent',
    status1 = if (sum(abun[Hour > window1] > 0)) 'present' else 'absent',
    status = if (status0 == 'absent' & status1 == 'absent') 'absent' else 'persistent',
    status = if (status0 == 'absent' & status1 == 'present') 'immigrant' else status,
    status = if (status0 == 'present' & status1 == 'absent') 'extinct' else status,
    ETA = if (status == 'immigrant') min(Hour[Hour > 97 & abun > 0]) else NA) %>%
  group_by(Pair, otu) %>%
  mutate(
    abun_pair = ifelse(!is.na(Pair) & Abx == 'Y', abun0[Abx == 'N'], NA),
    abun_pair = ifelse(!is.na(Pair) & Abx == 'N', abun0[Abx == 'Y'], abun_pair))

j <- x %>%
  group_by(Mouse, Abx, Pair, otu) %>%
  summarise(
    abun0 = sum(abun[Hour < window0]),
    abun1 = sum(abun[Hour >= window0 & Hour <= window1]),
    abun01 = abun0 + abun1,
    abun2 = sum(abun[Hour > window1]),
    status_treatment = if (abun01 == 0) 'absent' 
                         else if (abun0 > 0 & abun1 == 0) 'extinct' 
                           else 'persistent',
    ETD = if (status_treatment == 'extinct') max(Hour[Hour <= window0 & abun > 0]) 
            else if (status_treatment == 'persistent') Inf 
              else NA,
    status_recovery = if (abun01 == 0 & abun2 == 0) 'absent' 
               else if (abun1 == 0 & abun2 > 0) 'immigrant'
                 else if (abun01 > 0 & abun2 == 0) 'extinct'
                   else 'persistent',
    immigrant_type = if (status_recovery != 'immigrant') 'Not an immigrant'
                       else if (abun01 > 0) 'recovered'
                         else 'new colonist',
    ETA = if (status_recovery == 'immigrant') min(Hour[Hour > 97 & abun > 0]) 
            else NA) %>%
  group_by(Pair, otu) %>%
  mutate(
    abun_pair = ifelse(!is.na(Pair) & Abx == 'Y', abun01[Abx == 'N'], NA),
    abun_pair = ifelse(!is.na(Pair) & Abx == 'N', abun01[Abx == 'Y'], abun_pair))

p <- j %>% 
  filter(status_recovery == 'immigrant' & Abx == 'Y' & abun_pair > 0) %>%
  mutate(time_to_arrival = ETA - 120) %>%
  ggplot(aes(x = abun_pair, y = time_to_arrival)) + 
    geom_jitter(alpha = 0.25, color = 'red', height = 5, width = 0) + 
    scale_x_log10() + 
    stat_smooth(color = 'black', method = 'lm') +
    stat_smooth(color = 'black', lty = 3) +
    labs(x = "OTU relative abundance in cohoused control mouse", y = 'Hours before OTU arrival in Abx treated mouse') +
    th

@@

p

@@

**(ref:figPropPressure):** A test of the potential role of mass effects. The hypothesis is that, all else being equal, the abundance of a taxon in the regional pool (i.e., the cohoused mouse) is going to correlate positively with time to arrival (i.e., successful colonization) in the treated mouse. This hypothesis is supported. However, we can't be sure if the pattern is because the taxa that are more abundant taxa in the control mouse are simply better suited for the gut environment, and thus are faster to colonize the antibiotic-treated mouse. So, the analysis has it's limitations, but I still think it's worth noting.

@@

p <- x %>%
  ungroup() %>%
  left_join(j[, c('Mouse','otu','status_recovery','immigrant_type')], by = c('Mouse','otu')) %>%
  filter(!is.na(Pair) & abun > 0 & status_recovery != 'absent') %>%
  arrange(status_recovery) %>%
  mutate(status_recovery = ifelse(status_recovery == 'immigrant', immigrant_type, status_recovery)) %>%
  mutate(OTU_status = c('Extinct by end of Abx treatment (hour 120 at the latest); does not return during recovery period', 
                        'New colonist, arrived during recovery period (after hour 120)',
                        'Present throughout experiment',
                        'Extinct during end of Abx treatment (between Hours 63-120); returns during recovery period (after Hour 144')[match(as.numeric(factor(status_recovery, levels = sort(unique(status_recovery)))), c(1:4))],
         Pair = paste('Pair', Pair),
         Abx = ifelse(Abx == 'Y', 'Abx', 'No abx')) %>%
  ggplot(aes(x = as.factor(Hour), y = abun, fill = OTU_status)) +
    geom_bar(stat='identity', alpha = 0.8) +
    geom_rect(xmin = 0, xmax = 14, ymin = -100, ymax = 0, fill = 'red', alpha = 0.2) +
    facet_grid(Abx~Pair) +
    scale_fill_manual(name = '', values = c('#a6611a','#dfc27d','#80cdc1','#018571')) +
    #scale_fill_discrete(name = '') +
    labs(x = 'Hour: {0,4,8,12,16,20,23,25,39,49,63,73,97,120,144,168,192,216,240,312}', y = 'Rarefied abundance') +
    guides(fill=guide_legend(nrow=4,byrow=TRUE, position = 'bottom')) +
    theme_bw() +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      panel.background = element_blank(),
      axis.text.x=element_blank(),
      legend.position = 'bottom')

plot_it(p, name = 'figOTUAbunByAbxStatus', dir = image_dir, width = 6.5, height = 5)

p

@@

(ref:figOTUAbunByAbxStatus) Figure `r refs$figOTUAbunByAbxStatus`

`r pagebreak` ![](images\\figOTUAbunByAbxStatus.png)

**(ref:figOTUAbunByAbxStatus):** Rarefied OTU abundances, colored by OTU status. This figure is to compare mice treated by antibiotics to untreated mice.

x %>%
  ungroup() %>%
  left_join(j[, c('Mouse','otu','status_recovery','immigrant_type')], by = c('Mouse','otu')) %>%
  filter(Abx == 'Y' & abun > 0 & status_recovery != 'absent' & Mouse != 1) %>%
  arrange(status_recovery) %>%
  mutate(status_recovery = ifelse(status_recovery == 'immigrant', immigrant_type, status_recovery)) %>%
  mutate(OTU_status = c('Extinct by end of Abx treatment (hour 120 at the latest); does not return during recovery period', 
                        'New colonist, arrived during recovery period (after hour 120)',
                        'Present throughout experiment',
                        'Extinct during end of Abx treatment (between Hours 63-120); returns during recovery period (after Hour 144')[match(as.numeric(factor(status_recovery, levels = sort(unique(status_recovery)))), c(1:4))],
         Mouse = paste(ifelse(Cohoused == 'Y', 'Cohoused', 'Not cohoused'), "- Mouse", Mouse),
         Abx = ifelse(Abx == 'Y', 'Abx', 'No abx')) %>%
  ggplot(aes(x = as.factor(Hour), y = abun, fill = OTU_status)) +
    geom_bar(stat='identity', alpha = 0.8) +
    geom_rect(xmin = 0, xmax = 14, ymin = -100, ymax = 0, fill = 'red', alpha = 0.2) +
    facet_wrap(~Mouse, ncol = 5) +
    scale_fill_manual(name = '', values = c('#a6611a','#dfc27d','#80cdc1','#018571')) +
    #scale_fill_discrete(name = '') +
    labs(x = 'Hour: {0,4,8,12,16,20,23,25,39,49,63,73,97,120,144,168,192,216,240,312}', y = 'Rarefied abundance') +
    guides(fill=guide_legend(nrow=4,byrow=TRUE, position = 'bottom')) +
    theme_bw() +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      panel.background = element_blank(),
      axis.text.x=element_blank(),
      legend.position = 'bottom')

plot_it(p, name = 'figOTUAbunByCageStatus', dir = image_dir, width = 6.5, height = 5)

p

(ref:figOTUAbunByCageStatus) Figure `r refs$figOTUAbunByCageStatus`

`r pagebreak` ![](images\\figOTUAbunByCageStatus.png)

**(ref:figOTUAbunByCageStatus):** Rarefied OTU abundances, colored by OTU status. This figure is to compare mice in cohousing environments during the recovery process, with mice caged alone during the recovery process.

x %>%
  ungroup() %>%
  left_join(j[, c('Mouse','otu','status_recovery','immigrant_type')], by = c('Mouse','otu')) %>%
  filter(!is.na(Pair) & abun > 0 & status_recovery != 'absent') %>%
  arrange(status_recovery) %>%
  mutate(status_recovery = ifelse(status_recovery == 'immigrant', immigrant_type, status_recovery)) %>%
  mutate(OTU_status = c('Extinct by end of Abx treatment (hour 120 at the latest); does not return during recovery period', 
                        'New colonist, arrived during recovery period (after hour 120)',
                        'Present throughout experiment',
                        'Extinct during end of Abx treatment (between Hours 63-120); returns during recovery period (after Hour 144')[match(as.numeric(factor(status_recovery, levels = sort(unique(status_recovery)))), c(1:4))],
         Pair = paste('Pair', Pair),
         Abx = ifelse(Abx == 'Y', 'Abx', 'No abx')) %>%
  ggplot(aes(x = as.factor(Hour), fill = OTU_status)) +
    geom_bar(alpha = 0.8) +
    geom_rect(xmin = 0, xmax = 14, ymin = -100, ymax = 0, fill = 'red', alpha = 0.2) +
    facet_grid(Abx~Pair) +
    scale_fill_manual(name = '', values = c('#a6611a','#dfc27d','#80cdc1','#018571')) +
    #scale_fill_discrete(name = '') +
    labs(x = 'Hour: {0,4,8,12,16,20,23,25,39,49,63,73,97,120,144,168,192,216,240,312}', y = 'OTU richness') +
    guides(fill=guide_legend(nrow=4,byrow=TRUE, position = 'bottom')) +
    theme_bw() +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      panel.background = element_blank(),
      axis.text.x=element_blank(),
      legend.position = 'bottom')

plot_it(p, name = 'figOTURichByAbxStatus', dir = image_dir, width = 6.5, height = 5)

p

(ref:figOTURichByAbxStatus) Figure `r refs$figOTURichByAbxStatus`

`r pagebreak` ![](images\\figOTURichByAbxStatus.png)

**(ref:figOTURichByAbxStatus):** OTU richness from communities rarefied to 5000 seqeunces, colored by OTU status. This figure is to compare mice in cohousing environments during the recovery process, with mice caged alone during the recovery process.

x %>%
  ungroup() %>%
  left_join(j[, c('Mouse','otu','status_recovery','immigrant_type')], by = c('Mouse','otu')) %>%
  filter(Abx == 'Y' & abun > 0 & status_recovery != 'absent' & Mouse != 1) %>%
  arrange(status_recovery) %>%
  mutate(status_recovery = ifelse(status_recovery == 'immigrant', immigrant_type, status_recovery)) %>%
  mutate(OTU_status = c('Extinct by end of Abx treatment (hour 120 at the latest); does not return during recovery period', 
                        'New colonist, arrived during recovery period (after hour 120)',
                        'Present throughout experiment',
                        'Extinct during end of Abx treatment (between Hours 63-120); returns during recovery period (after Hour 144')[match(as.numeric(factor(status_recovery, levels = sort(unique(status_recovery)))), c(1:4))],
         Mouse = paste(ifelse(Cohoused == 'Y', 'Cohoused', 'Not cohoused'), "- Mouse", Mouse),
         Abx = ifelse(Abx == 'Y', 'Abx', 'No abx')) %>%
  ggplot(aes(x = as.factor(Hour), fill = OTU_status)) +
    geom_bar(alpha = 0.8) +
    geom_rect(xmin = 0, xmax = 14, ymin = -100, ymax = 0, fill = 'red', alpha = 0.2) +
    facet_wrap(~Mouse, ncol = 5) +
    scale_fill_manual(name = '', values = c('#a6611a','#dfc27d','#80cdc1','#018571')) +
    #scale_fill_discrete(name = '') +
    labs(x = 'Hour: {0,4,8,12,16,20,23,25,39,49,63,73,97,120,144,168,192,216,240,312}', y = 'OTU richness') +
    guides(fill=guide_legend(nrow=4,byrow=TRUE, position = 'bottom')) +
    theme_bw() +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      panel.background = element_blank(),
      axis.text.x=element_blank(),
      legend.position = 'bottom')

plot_it(p, name = 'figOTURichByCageStatus', dir = image_dir, width = 6.5, height = 5)

p

(ref:figOTURichByCageStatus) Figure `r refs$figOTURichByCageStatus`

`r pagebreak` ![](images\\figOTURichByCageStatus.png)

**(ref:figOTURichByCageStatus):** OTU richness from communities rarefied to 5000 seqeunces, colored by OTU status. This figure is to compare mice in cohousing environments during the recovery process, with mice caged alone during the recovery process.

@@

p <- x %>%
  left_join(tax) %>%
  filter(otu %in% j$otu) %>%
  ungroup() %>%
  mutate(
    label = paste(Family, 'OTU', as.numeric(as.factor(otu))),
    Abx = ifelse(Abx == 'N', 'No abx', 'Abx')) %>%
  left_join(j) %>%
  ggplot(aes(x = Hour, y = abun + 1, color = Abx)) + 
  annotate(geom = 'rect', xmin = 1, xmax = 120, ymin = 0, ymax = Inf, alpha = 0.3, fill = 'red') + 
  geom_point() + 
  stat_smooth() +
  #geom_text(label = "Antibiotics", x = 60, aes(y = max(abun))) +
  #geom_text(label = "Recovery", x = 216, aes(y = max(abun))) +
  scale_y_log10() +
  scale_color_manual(values = c('red','black'), name = '') +
  facet_wrap(~label) + 
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.background = element_blank(),
    legend.position = 'bottom')

@@

p
```

@@

`r pagebreak` ![](images\\figMostAffectedOTUs.png)

**(ref:figMostAffectedOTUs):** Trends in abundances of 12 taxa that differ significantly between antibiotic treated mice and control mice. The translucent red square shows the window of antibiotic treatment.
